Radio Emission from Red-Giant Hot Jupiters

### 2015-09-28 ms.tex
hot jupiters => jupiters where the description is also true for radio emission of any kind

An additional factor that affects detectability is source confusion, the imaging noise associated with unresolved interloping radio sources in an observation.
Source confusion is characterized by the standard deviation $\sigma_c$ in the surface brightness of an image due to one or more unresolved sources in the beam solid angle (for a review, see \cite{Condon1974,Condon2012}).
To estimate $\sigma_c$, reported in Table~\ref{tab:sens}, we use the relations provided by \cite{condon+ransom2016}, reproduced below:
%%%
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
%%%
As Table~\ref{tab:sens} illustrates, source confusion is likely to exceed the 2 $\mu$Jy detection limit of a typical hot Jupiter source for most
telescopes under consideration.  However, source confusion can be mitigated through differencing observations that cancel out background sources.
Potential axes for such cancellation are polarization and time.
The source confusion limit is dominated by active galactic nuclei, which have low polarization fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time.
Therefore both the circularly polarized nature of the cyclotron emission from hot jupiters and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate the source confusion limit, allowing hot jupiters, like other time-variable sources, to be detected below the imaging confusion limit.
We therefore expect confusion not to be a strong limitation to the detectability of hot jupiters.

@@

% SKA
The low frequency component of the Square Kilometer Array (SKA), recently rebaselined to have 0.4~km$^2$ collecting area in 130,000 elements with baselines extending to 65~km, will operate from 50-350~MHz.\footnote{\url{https://www.skatelescope.org/news/worlds-largest-radio-telescope-near-construction/}}
We estimate the noise in the lower half of the band ($\nu<200~\rm MHz$), appropriate for lower mass ($\lesssim 10 M_{\rm J}$) separately from the full band, which could be used to detect $\sim 30 M_{\rm J}$ hot Jupiters.
We again adopt the \cite{Rogers+Bowman2008} relation for \Tsys\ (Equation~\ref{eq:Tsky}), and assume $\Trx \approx 60~\rm K$.
We caution that these assumptions are optimistic, and the time estimates could in reality be a factor of a few worse.

@@

The noise estimates shown in Table \ref{tab:sens}, together with Figure \ref{fig:observability}, indicate that, at the frequencies relevant for the detection of radio emission from Jupiter-size planets, current instrument sensitivities are generally too low.
%On the other hand, for more massive exoplanets ($M_p\sim 10 M_{\rm J}$) where the bulk of the radio emission would exist above 50~MHz, there is hope to detect radio emission from nearby systems with LOFAR-HBA or LOBO.
%LOFAR-HBA has reasonable sensitivity to $M_p =10~M_{\rm J}$ planets at 5 AU around the few closest systems as well as those around further stars but orbiting closer to the stars. 
On the other hand, for more massive exoplanets ($M_p\sim 10 M_{\rm J}$) where the bulk of the radio emission would exist above 50~MHz, low frequency component of SKA will be able to detect the signal within one day from planets with $M_p \sim 10~M_{\rm J}$ at $\sim$5~AU around M stars within $\sim $100~pc from Earth, as well as those around even further stars but orbiting closer to the stars. 
VLA-LOBO could also work for a few nearby systems.

>>>>

An additional factor that affects detectability is source confusion, the imaging noise associated with unresolved interloping radio sources in an observation.
Source confusion is characterized by the standard deviation $\sigma_c$ in the surface brightness of an image due to one or more unresolved sources in the beam solid angle (for a review, see \cite{Condon1974,Condon2012}).
To estimate $\sigma_c$, reported in Table~\ref{tab:sens}, we use the relations provided by \cite{condon+ransom2016}, reproduced below:
%%%
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
%%%
As Table~\ref{tab:sens} illustrates, source confusion is likely to exceed the 2 $\mu$Jy detection limit of a typical RGHJ source for most
telescopes under consideration.  However, source confusion can be mitigated through differencing observations that cancel out background sources.
Potential axes for such cancellation are polarization and time.
The source confusion limit is dominated by active galactic nuclei, which have low polarization fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time.
Therefore both the circularly polarized nature of the cyclotron emission from planets and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate the source confusion limit, allowing planetary radio emission, like other time-variable sources, to be detected below the imaging confusion limit.
We therefore expect confusion not to be a strong limitation to the detectability of radio emission from RGHJs.

@@

% SKA
The low frequency component of the Square Kilometer Array (SKA), recently rebaselined to have 0.4~km$^2$ collecting area in 130,000 elements with baselines extending to 65~km, will operate from 50-350~MHz.\footnote{\url{https://www.skatelescope.org/news/worlds-largest-radio-telescope-near-construction/}}
We estimate the noise in the lower half of the band ($\nu<200~\rm MHz$), appropriate for lower mass ($\lesssim 10 M_{\rm J}$) separately from the full band, which could be used to detect $\sim 30 M_{\rm J}$ Jupiters.
We again adopt the \cite{Rogers+Bowman2008} relation for \Tsys\ (Equation~\ref{eq:Tsky}), and assume $\Trx \approx 60~\rm K$.
We caution that these assumptions are optimistic, and the time estimates could in reality be a factor of a few worse.

@@

The noise estimates shown in Table \ref{tab:sens}, together with Figure \ref{fig:observability}, indicate that, at the frequencies relevant for the detection of radio emission from Jupiter-size planets, current instrument sensitivities are generally too low.
%On the other hand, for more massive exoplanets ($M_p\sim 10 M_{\rm J}$) where the bulk of the radio emission would exist above 50~MHz, there is hope to detect radio emission from nearby systems with LOFAR-HBA or LOBO.
%LOFAR-HBA has reasonable sensitivity to $M_p =10~M_{\rm J}$ planets at 5 AU around the few closest systems as well as those around further stars but orbiting closer to the stars. 
On the other hand, for more massive exoplanets ($M_p\sim 10 M_{\rm J}$) where the bulk of the radio emission would exist above 50~MHz, low frequency component of SKA will be able to detect the signal within one day from planets with $M_p \sim 10~M_{\rm J}$ at $\sim$1~AU and at $\sim$5~AU within $\sim $200~pc and $\sim $100~pc from Earth, respectively.  
VLA-LOBO could also work for a few nearby systems.
